% !TeX root = ../../../main.tex

As briefly introduced in \cref{sec:qcd/dglap}, a central element of a \pdf fit
consists in \dglap evolution.
%
This will be used in two contexts: 
\begin{enumerate}[label=\roman*.]
  \item upgrading grids to the so called \fktab{}s (cf. \cref{ch:pine}), to
    obtain theory predictions at many different scales (matching the
    experimental results), starting from the \pdf candidate being minimized at
    a single fitting scale\footnote{
      Or, equivalently, evolving online during the fit at all those scales.
    }
  \item evolving the result of the fit, to distribute \pdf grids for the users
    already at all the potentially useful scales (interpolating among a finite
    sets of values)
\end{enumerate}

While working on new \pdf fits and related issues, a certain set of limitations
of the existing \dglap evolution libraries has been collected, and for this
reason we decided to write a new \qcd{} evolution library that matches the
requirements of a \pdf fitting collaboration, and similar fits as well.

We called it \acrfull{eko}, because of the main focus is not to compute
evolution kernel operators (EKO) which are independent of the initial boundary
conditions but only depend on the given theory settings.
In such manner, the operators can thus be computed once, stored on disk and
then reused in the actual application. This method can lead to a significant
speed-up when \pdfs are repeatedly being evolved, as it is customary in \pdf{}
fits.
This approach has been introduced by
\fk{}~\cite{Ball:2008by,Ball:2010de,DelDebbio:2007ee} and it is further
developed here.
%
Furthermore, we decided to solve the evolution equations~ in Mellin space (cf.\
\cref{sec:theory:mellin}) to allow for simpler solution algorithms (both
iterative and approximated).
Yet, \eko provides result in momentum fraction space (cf.\
\cref{sec:theory:interpolation}) to allow an easy interface with existing
codes.

\eko{} is open-source, allowing easy interaction with users
and developers.
The project comes with a clean, modular, and maintainable codebase that guarantees
easy inspection and ensures it is future-proof.
We provide both a user and a developer documentation. So, not only a user
manual, but even the internal documentation is published, with an effort to make
it as clear as possible.

\eko{} currently implements the leading order (\lo{}),
next-to-leading order (\nlo{}) and next-to-next-to-leading order (\nnlo{})
solutions~\cite{Vogt:2004mw,Moch:2004pa,Blumlein:2021enk}.
However, it is organized in such a way that the addition of higher order corrections,
such as the next-to-next-to-next-to-leading order (\nnnlo{})~\cite{Moch:2021qrk},
can be achieved with relative ease.
This accuracy is needed to match the precision in the determination of the
matrix elements for several processes at LHC (see e.g.\ \cite{Duhr:2021vwj} and
references therein).
We also expose the associated variations of the various scales.

\eko{} correctly treats intrinsic heavy quark distributions,
required for studies of the heavy quark content of the nucleon~\cite{Ball:2022qks}.
While the treatment of intrinsic distributions in the evolution equations is
mathematically simple, as they decouple in a specific basis, their integration
into the full solution, including matching conditions, is non-trivial.
We also implement backward evolution, again including the non-trivial
handling of matching conditions.

\eko{} is another corner stone in a suite of tools that aims to
provide new handles to the theory predictions in the \pdf{} fitting
procedure. To obtain the theory predictions in a typical fitting procedure,
first, one needs to evolve a candidate \pdf{} up to the process scale,
and then convolute it with the respective coefficient function.
The process specific coefficient function can be stored in the
\pineappl{} format~\cite{Carrazza_2020,christopher_schwan_2022_5846421}.
\eko{} is interfaced with \pineappl{} to produce
interpolation grids that can be directly used in a \pdf{} fit,
avoiding to do the evolution on the fly, but beforehand.
\eko{} is also powering \yadism{}~\cite{yadism} a new \dis{}
coefficient function library.

\eko{} adopts Python as a programming language opting for a high-level language
which is easy to understand for newcomers.
In particular, with the advent of Data Science and Machine Learning, Python has
become the language of choice for many scientific applications, mainly driven
by the large availability of packages and frameworks.
We decided to write a code that can be used by everyone who needs \qcd{}
evolution, and to make it possible for applications that are not supported yet
to be built on top of the provided tools and ingredients.
For this reason the code is developed mainly as a library, that contains
physics, math, and algorithmic tools, such as those needed for managing or
storing the computed operators. As an example we also expose a runner, making
use of the built library to deliver an actual evolution application. 
We apply modern best practices for software development, such as automated
tests, Continuous Integration (CI), and Continuous Deployment (CD), to ensure a
high quality of coding standard and a routinely checked code basis.

\paragraph{References} The open-source repository is available at:
\begin{center}
\url{https://github.com/N3PDF/eko}
\end{center}
The \eko documentation contains the full \api reference, tutorials to get
started with \eko calculations, and far more complete discussion about the
underlying theory and numerical techniques adopted.
It can be accessed at:
\begin{center}
\url{https://eko.readthedocs.io/en/latest/}
\end{center}
This document is also regularly updated and extended upon the
implementation of new features.
