We do not attempt to give a full review of the underlying theory
here as it is known since a long time and discussed extensive elsewhere
(see e.g.\ \cite{Peskin:1995ev,Ellis:1996mzs} and references therein).
We refer the interested reader to the specific references given in the following and to
the accompanying online documentation where instead we give a detailed
overview. All sections in the following have an equivalent section in
the online documentation. Also the respective code implementations of the
various ingredients contain relevant information and are also accessible
in the documentation via the API section.

The central equations that \eko{} is solving are the
Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (\dglap) evolution
equations~\cite{Altarelli:1977zs,Gribov:1972ri,Dokshitzer:1977sg} given by
\begin{equation}
    \muF^2 \dv{\vb f}{\muF^2}{}(x,\muF^2) = \vb P (a_s(\muR^2),\muF^2) \otimes \vb f(\muF^2)
    \label{eq:eko/dglap}
\end{equation}
where $\vb f(x,\muF^2)$ is a vector of \pdf{}s over flavor space with $x$ the
momentum fraction and $\muF^2$ the factorization scale.
The main ingredients to \cref{eq:eko/dglap} are the Altarelli-Parisi splitting
functions $\vb P(a_s(\muR^2),x,\muF^2)$~\cite{Moch:2004pa,Vogt:2004mw}, which
are matrices over the flavor space.
Finally, $\otimes$ denotes the multiplicative (or Mellin) convolution.

The splitting functions $\vb P(a_s(\muR^2),x,\muF^2)$ expose a perturbative
expansion in the strong coupling $a_s(\mu_R^2)$:
\begin{equation}
    \vb P(a_s(\muR^2),x,\muF^2) = a_s(\muR^2) \vb P^{(0)}(x,\muF^2)
    + \qty[a_s(\muR^2)]^2 \vb P^{(1)}(x,\muF^2)
    + \qty[a_s(\muR^2)]^3 \vb P^{(2)}(x,\muF^2)
    + \ldots
\end{equation}
which is currently known at \nnlo{}~\cite{Moch:2004pa,Vogt:2004mw,Blumlein:2021enk} and is under
investigation for \nnnlo{}~\cite{Moch:2021qrk}.
In a first step, the renormalization scale $\muR$ and the factorization scale
$\muF$ can be assumed to be equal $\muR = \muF$ and the renormalization scale
dependence can be restored later on. The variation of the ratio $\muR/\muF$ can
be considered as an estimated to missing higher order
uncertainties (\mhou{})~\cite{AbdulKhalek:2019ihb}.

In order to solve \cref{eq:eko/dglap} a series of steps has to be taken, and we
highlight these steps in the following sections.

\subsection{Mellin space}
\label{sec:theory:mellin}
\input{sections/chapters/eko/theory/mellin.tex}

\subsection{Interpolation}
\label{sec:theory:interpolation}
\input{sections/chapters/eko/theory/interpolation.tex}

\subsection{Strong coupling}
\label{sec:theory:coupling}
\input{sections/chapters/eko/theory/coupling.tex}

\subsection{Flavor space}
\label{sec:theory:flavor}
\input{sections/chapters/eko/theory/flavor.tex}

\subsection{Solution Strategies}
\label{sec:theory:solutions}
\input{sections/chapters/eko/theory/solutions.tex}

\subsection{Matching at Thresholds}
\label{sec:theory:matching}
\input{sections/chapters/eko/theory/matching.tex}

\subsection{Running Quark Masses}
\label{sec:theory:msbarmass}
\input{sections/chapters/eko/theory/msbar-masses.tex}
